|

### 2023-09-07 manuscript/manuscript_1.0.Rmd
New version of the manuscript

In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974] or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems in which individuals 
exhibit badges of status, where a signal conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
Such signals must be arbitrary, meaning the signals could potentially be 
produced by high and low quality individuals alike. 
Generally signals can be evolutionarily stable whenever low-quality individuals 
pay a higher fitness cost for carrying the signal such that it is not 
in the interest of low-quality individuals to fake quality by producing a large badge.
Signal costs inversely proportional to quality may come about as a result of the
production costs of the signal, in which case the signal works like a handicap
[@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990;@zahavi_Mate_1975]. 
Alternatively, the costs can come from 
the aggressive reaction of receivers when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. For example in paper 
wasps, subordinate individuals with experimentally manipulated dominant-like facial
patterns received more aggression from the dominant[@tibbetts_Socially_2004]. 
In this case, costs are socially rather than developmentally mediated and are 
triggered by the violation of the convention, although how the convention
is established in the first place is unclear.

@@

The action of natural selection and genetic drift on the learning parameters 
do not override the conditions for the evolutionary branching 
[@dieckmann_Dynamical_1996] described earlier. 
Given how crucial naive behaviour is for the branching event 
described earlier, we tested whether evolutionary processes could lead 
key parameters of the learning process toward
values that prevented the genetic variation to build up. 
Figure \ref{fig:learnEvol} shows the dynamics from a set of simulations where we 
allow for the coevolution of the reaction norm as well as key parameters of
the learning module, namely the speed of learning ($A$) and the behavioural 
tendency expressed by individuals prior to learning. Top panels in 
figure \ref{fig:learnEvol} show that there is diversification process in this 
coevolutionary scenario as well. That is because, as is shown in the middle 
panels (B) of the same figure, the learning parameters stay within the range 
necessary for the branching event to take place. The speed of learning 
(Fig. \ref{fig:learnEvol} B.1) stays well above zero throughout the simulation. 
In contrast, the initial behavioral tendency maintain values around zero. 
Thus, evolutionary processes acting on the learning parameters 
did not override the conditions necessary for frequency-dependent selection to boost 
genetic variation.

@@

Originally, @rohwer_Social_1975a proposed the idea that certain phenotypic 
traits could be used by animals as status signals or signals of quality. 
This idea has been tested repeatedly on the dark patches of some bird species, 
such as the bibs exhibited by species in various families, particularly sparrows.
So, are bibs real signals of quality? if so are they handicaps or 
badge of status? From an evolutionary perspective it is unclear how the stability of 
status signals can be maintained. One option is that honesty is 
maintained by the cost of signal production; that is, when the signal works as a 
handicap [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993]. 
Alternatively, if the signal does not carry production costs inversely 
proportional to quality it may work as a convention.
In this later case, honesty is presumably maintained by the aggressive 
reaction of receivers when the convention is 
broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model, portraits 
both kinds of signals, although with some nuances. When we assume the handicap 
principle, simulations result in the evolution of honest signals mediated by the
learned response. When we assume cost-free signal, under certain conditions the
learned response mediated the emergence of phenotypic variation in the signal. 
This variation, particularly in the mid-range of the distribution facilitates 
the establishment of a convention, in this range individuals respond 
appropriately to the trait of their peer. The house sparrow (*Passer domesticus*) 
has been a text-book example of badges of status. According to the prevailing narrative the bib size in males is a signal of dominance rank and quality.
However, a recent meta-analysis called into question 
this narrative by showing that the effect size of the association between bib 
size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018a]. 
A simple correlation between bib size and dominance rank
(and quality) is expected under the handicap principle, as it is shown here and 
elsewhere [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993].
However, the branching event presented here shows that conventions can 
emerge in a less straightforward way when mediated by learning processes. These 
nuances could potentially make sense of seemingly contradictory evidence on the 
correlation between plumage traits and quality. For example, conventions in our
simulations arose only in an intermediate range of the signalling trait, yet 
empirical studies seeing for correlations between phenotypic traits and quality 
are often performed across the whole range of variation.

The focus of theoretical and empirical work on communication systems, 
and particularly badges of status, is often explaining the presence and 
absence of certain morphological traits within populations. However, the 
morphological traits hypothesized to play a role in communication systems 
regularly present interesting patters at the phylogeographic
level. One example of this is the leapfrog pattern [@remsen_High_1984],
whereas a certain morphological trait alternates its presence and 
absence in a set of geographically adjunct populations. Patterns of 
molecular variation along geographic clines suggest that natural selection 
is involved in the emergence of these patterns 
[@cadena_Testing_2011;@cadena_Evolutionary_2007]. However, it is unclear what 
type of ecological process is behind these selective regimes. 
Given the wide variation in outcomes found in our model, that depend on both 
stochasticity and cognitive parameters, we think variation in cognition and 
learning in particular could provide some explanatory power in this respect. 
One first step, for example is to evaluate the way individuals in different 
populations respond to novel traits [@avendano_Territorial_2021].

>>>>

In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974] or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems in which individuals 
exhibit badges of status, where a signal conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975]. 
Such signals tend to be arbitrary, meaning the signals could potentially be 
produced by high- and low-quality individuals alike. This symmetry opens up the 
possibility of dishonest signalling and the question of evolutionary stability. 
Signals can be evolutionarily stable whenever low-quality individuals 
pay a higher fitness cost for carrying the signal such that it is not 
in the interest of low-quality individuals to fake quality by producing a large badge.
Signal costs inversely proportional to quality may come about as a result of the
production costs of the signal, in which case the signal works like a handicap
[@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990;@zahavi_Mate_1975]. 
Alternatively, the costs can come from 
the aggressive reaction of receivers when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. For example in paper 
wasps, subordinate individuals with experimentally manipulated dominant-like facial
patterns received more aggression from the dominant[@tibbetts_Socially_2004]. 
In this case, costs are socially rather than developmentally mediated and are 
triggered by the violation of the convention, although how the convention
is established in the first place is unclear.

@@

The action of natural selection and genetic drift on the learning parameters 
do not override the conditions for the evolutionary branching 
[@dieckmann_Dynamical_1996] described earlier. 
Given how crucial naive behaviour is for the branching event 
described earlier, we tested whether evolutionary processes could lead 
key parameters of the learning process toward
values that prevented the genetic variation to build up. 
Figure \ref{fig:learnEvol} shows the dynamics from a set of simulations where we 
allow for the coevolution of the reaction norm as well as key parameters of
the learning module, namely the speed of learning ($A$) and the behavioural 
tendency expressed by individuals prior to learning. Top panels in 
figure \ref{fig:learnEvol} show that there is diversification process in this 
coevolutionary scenario as well. That is because, as is shown in the middle 
panels (B) of the same figure, the learning parameters stay within the range 
necessary for the branching event to take place. The speed of learning 
(Fig. \ref{fig:learnEvol} B.1) stays well above zero throughout the simulation. 
In contrast, the initial behavioural tendency maintain values around zero. 
Thus, evolutionary processes acting on the learning parameters 
did not override the conditions necessary for frequency-dependent selection to boost 
genetic and phenotypic variation.

@@

Originally, @rohwer_Social_1975 proposed the idea that certain phenotypic 
traits could be used by animals as status signals or signals of quality. 
This idea has been tested repeatedly on the dark patches of some bird species, 
such as the bibs exhibited by species in various families, particularly sparrows.
So, are bibs real signals of quality? if so are they handicaps or 
badge of status? From an evolutionary perspective it is unclear how the stability of 
status signals can be maintained. One option is that honesty is 
maintained by the cost of signal production; that is, when the signal works as a 
handicap [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993]. 
Alternatively, if the signal does not carry production costs inversely 
proportional to quality it may work as a convention.
In this later case, honesty is presumably maintained by the aggressive 
reaction of receivers when the convention is 
broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model, captures 
both kinds of signals, although with some nuances. When we assume the handicap 
principle, simulations result in the evolution of honest signals mediated by the
learned response. When we assume cost-free signal, under certain conditions the
learned response mediated the emergence of phenotypic variation in the signal. 
This variation, particularly in the mid-range of the distribution facilitates 
the establishment of a convention, in this range individuals respond 
appropriately to the trait of their peer. The house sparrow (*Passer domesticus*) 
has been a text-book example of badges of status. According to the prevailing narrative the bib size in males is a signal of dominance rank and quality.
However, a recent meta-analysis called into question 
this narrative by showing that the effect size of the association between bib 
size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018]. 
A simple correlation between bib size and dominance rank
(and quality) is expected under the handicap principle, as it is shown here and 
elsewhere [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993].
However, the branching event presented here shows that conventions can 
emerge in a less straightforward way when mediated by learning processes. These 
nuances could potentially make sense of seemingly contradictory evidence on the 
correlation between plumage traits and quality. For example, conventions in our
simulations arose only in an intermediate range of the signalling trait, yet 
empirical studies seeing for correlations between phenotypic traits and quality 
are often performed across the whole range of variation.

The focus of theoretical and empirical work on communication systems, 
and particularly badges of status, is often explaining the presence and 
absence of certain morphological traits within populations. Phenotypic traits
hypothesized to play a role in communication systems regularly exhibit 
patterns of geographic variation within species that have long called 
for explanations. One example of this is the leapfrog pattern [@remsen_High_1984],
whereby a certain morphological trait alternates its presence and 
absence in a set of geographically adjunct populations, likely as 
a consequence of selection [@cadena_Testing_2011;@marquez_Divergence_2020]. 
However, it is unclear what 
type of ecological process is behind these selective regimes. 
Given the wide variation in outcomes found in our model, which depend on both
stochasticity and cognitive parameters which may vary in space owing to various
processes, we speculate that variation in cognition and learning could 
provide some explanatory power in this respect. A first step to assess this 
hypothesis would be to evaluate the way individuals in different populations 
respond to novel traits [*e.g.* @avendano_Territorial_2021].
